---
title: "KRAFTWER KORP SALES ANALYSIS"
theme: default
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
---

# **Problem Statement**
Kraftwerk Korp is Atlas AI's new customer. They have 500 stores spread across Nairobi selling 3 different products: Bits, Bobs and Widgets. Kraftwerk has shared with you as a data analyst their sales data which include the location of each store and total sales volume for each product and expect you as a data analyst to help them answer two key questions:


1.What is the relationship between store performance and location?  
2.Based on that relationship, where should Kraftwerk build more stores?  

# **Data Analysis Guide**
We will use the Google Data Analytics framework to guide us through the analysis:


1.**Ask** - **Business Challenge/Objective/Question**

  The key business questions that Kraftwerk wants to answer include:

i)What are the top 5 neighborhoods in terms of revenue?
  
ii)What are the top 5 stores in terms of sales quantity?
  
iii)What is the best selling item in terms of both sales quanity and revenue?
  
iv)Which part of Nairobi has the highest concentration of stores?

2.**Prepare** - **Data generation, collection, storage and management**

  We will load the sales data into R studio session and explore it to understand its structure and properties.

3.**Process** - **Data cleaning and transformation**

  The data cleaning in this case will involve:

-inspecting for NAs/missing values

-formatting column names

-checking for duplicate records

 Data transformation will entail:

-calculating revenue per product

-sorting sales and revenue data

-calculating total revenue and sales for all the products

4.**Analyze** - **Data exploration, visualization and analysis**

  The analysis phase will involve creating static visualizations such as bar plots, box   plots and drawing conclusions.

5.**Share** - **Communicating and interpreting results**

  We will use R shiny tool to help bring the data to life and answer the key business questions.

6.**Act** - **Putting insights to work**

  Finally, we will make data-driven recommendations to Kraftwerk on store performance and expansion strategies

# **Part 1: General EDA - Getting to Know the Data**
## **1.1 Importing Required Packages**
If the required libraries are not installed in your R studio session, the following script will install and load the libraries
```{r libraries, echo=TRUE,warning=FALSE,error=FALSE,message=FALSE}
#script installs required packages if not installed in the R session
if(!require(rmarkdown)) install.packages("rmarkdown")
if(!require(tinytex)) install.packages("tinytex")
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(skimr)) install.packages("skimr")
if(!require(readr)) install.packages("readr")
if(!require(extrafont)) install.packages("extrafont")
if(!require(extrafont)) install.packages("ggthemes")
if(!require(ggtext)) install.packages("ggtext")
if(!require(ggtext)) install.packages("leaflet")


#load the packages
library(rmarkdown)
library(tinytex)
library(tidyverse)
library(skimr)
library(readr)
library(extrafont)
library(ggthemes)
library(ggtext)
library(leaflet)

#font_import()
#loadfonts(device = "win")

```

## **1.2 Data Preparation**
### **1.2.1 Load the data**
```{r dataset load,echo=TRUE,warning=FALSE,error=FALSE,message=FALSE}

data_path = "./www/Kraftwerk_korp_sales.csv"

sales_data = read_csv(data_path)

#script to have the data overview
head(sales_data)
```

### **1.2.2 Inspect the data**

The sales data has 501 rows, 8 columns and no missing values.
```{r dataset structure, echo=TRUE}
#creating function to understand the data structure
data_structure <- function(df){
  
  #column data types
  str(df)
  
  #summary statistics mean,median,max,min
  summary(df)
  
  #checks for missing values,sd,p0,p25,p50,p75,p100
  skim(df)
}

#applying the function to the data
data_structure(sales_data)
```

### **1.2.1 Checking for duplicates**

There are no duplicate data entries in the dataset
```{r check for duplicates, echo=TRUE}
sum(duplicated(sales_data))
```

## **1.3 Data Cleaning and Transformation**
Renaming the columns
```{r rename columns, echo=TRUE}
#renaming columns
sales_data <- sales_data %>%
  rename(
    'shop_location' = neighborhood,
    'lon' = y,
    'lat' = x
  )
head(sales_data)
```

Calculating the products' revenue for each store
```{r products revenue, echo=TRUE}
sales_revenue_data <- sales_data %>%
  mutate(
    bits_revenue = bits_qty*100,
    bobs_revenue = bobs_bqty*150,
    widgets_revenue = widgets_qty*300,
    total_revenue = bits_revenue+bobs_revenue+widgets_revenue,
    total_sales =  bits_qty+bobs_bqty+widgets_qty
    
  )
head(sales_revenue_data)
```
Saving the processed and cleaned sales data
```{r}
write.csv(sales_revenue_data,"./www/processed_sales_data.csv",row.names = FALSE)
```

calculating cumulative sales by neighborhood
```{r neighborhood revenue, echo=TRUE}
sales_revenue_neighborhood <- sales_revenue_data %>%
  group_by(shop_location) %>%
  summarise(
    neighborhood_revenue = sum(total_revenue)
  )
head(sales_revenue_neighborhood)
```


# **Part 2: General EDA - Answering Business Questions**

We will set a plotting theme for our plots
```{r plotting custom theme, echo=TRUE}
#setting theme
f1<-"Comic Sans MS"

custom_theme<-function()
{
  #setting font
Font<-"Comic Sans MS"

theme_economist() %+replace%
  
  theme(
    ########grid elements########
    panel.grid.major = element_blank(),    #strip major gridlines
    panel.grid.minor = element_blank(),    #strip minor gridlines
     axis.ticks = element_blank(),          #strip axis ticks
    axis.line = element_blank(),
    ######text elements###########
    plot.title = element_markdown(             #title
                   family = Font,            #set font family
                   size = 13,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = 0,                #left align
                   vjust = 2),               #raise slightly
      
      plot.subtitle = element_markdown(          #subtitle
                   family = Font,            #font family
                   size = 12,hjust = 0),               #font size
      
      plot.caption = element_text(           #caption
                   family = Font,            #font family
                   size = 8.7,
                   hjust = 0
                   ),               #right align
      
      axis.title = element_text( family = Font,size = 11),               #font size
      
      axis.text = element_text( family = Font,size = 11),               #font size
    
      legend.text = element_text( family = Font,size = 11),
      legend.title = element_text( family = Font,size = 11),
      legend.position = "right",
      plot.background = element_rect(fill = "white", color = NA)

    
    ###########plot#################
    
  )
}
```

### **Q1 - What are the top 5 neighborhoods in terms of revenue?**
```{r Q1, echo=TRUE}
#filtering top 5 neighborhoods in terms of revenue
 top5_neighborhoods <-  sales_revenue_neighborhood %>% 
  arrange(desc(neighborhood_revenue)) %>% 
  top_n(5)

head(top5_neighborhoods)

```
Plotting top 5 neighborhoods in terms of revenue
```{r top 5 neighborhood revenue, echo=TRUE,fig.height=6,fig.width=14,fig.align='left',warning=FALSE}
top5_neighborhoods %>% 
  ggplot(aes(x=reorder(shop_location,neighborhood_revenue),y=neighborhood_revenue))+
  geom_col(fill="#ffc107")+
  labs(title ="Top 5 neighborhoods in terms of revenue",y="",x="Neighborhood")+
  geom_text(aes(label=scales::comma(neighborhood_revenue,prefix = "Ksh")),color="black",
            hjust=0.6,vjust=0.5,family=f1,size=2.8)+
  coord_flip()+
  custom_theme()+
  theme(
  axis.text.x = element_blank()
  )
```

### **Q2 - What are the top 5 stores in terms of sales quantity?**

```{r Q2, echo=TRUE}
#filtering top 5 stores in terms of sales quantity
top5_stores <-  sales_revenue_data %>% 
  select(shop_location,shop_name,total_sales) %>%
  arrange(desc(total_sales)) %>% 
  top_n(5)

#inspecting data
head(top5_stores)
```
Plotting top 5 stores in terms of sales quantity
```{r top 5 stores sales quantity, echo=TRUE,fig.height=6,fig.width=14,fig.align='left',warning=FALSE}
top5_stores %>% 
  ggplot(aes(x=reorder(shop_name,total_sales),y=total_sales))+
  geom_col(fill="#ffc107")+
  labs(title ="Top 5 stores in terms of sales quantity",y="",x="Shop name")+
  geom_text(aes(label=scales::comma(total_sales)),color="black",
            hjust=0.6,vjust=0.5,family=f1,size=2.8)+
  coord_flip()+
  custom_theme()+
  theme(
  axis.text.x = element_blank()
  )
```

### **Q3 - What is the best-selling item in terms of both sales quantity and revenue?**
```{r q3, echo=TRUE}
#converting sales_revenue data from wide to long format
#grouping by product to calculate the total sales volume and revenue for each product
sales_revenue_data_long <- sales_revenue_data %>% 
  select(bits_qty:widgets_revenue) %>%
  gather(product, amount,bits_qty:widgets_revenue)%>%
  group_by(product) %>%
  summarise(
    total_amount = sum(amount)
  )
sales_revenue_data_long$type <- c("sales volume","revenue","sales volume","revenue","sales volume","revenue")

sales_revenue_data_long$product <-c("bits","bits","bobs","bobs","widgets","widgets")
#inspecting the data
head(sales_revenue_data_long)
```
Plotting the best-selling items in terms of both sales quantity and revenue
```{r best selling items plot, echo=TRUE,fig.height=6,fig.width=14,fig.align='left',warning=FALSE}
sales_revenue_data_long %>% 
  ggplot(aes(x=product,y=total_amount,fill=type))+
  geom_bar(position = "dodge",stat = "identity")+
  geom_text(aes(label=scales::comma(total_amount)),family = f1,size=2.4,hjust=0.4,vjust=0.5,color="black",
             position=position_dodge(width=0.9),face="bold")+
  labs(title = "Best-selling items in terms of both sales quantity and revenue")+
  scale_fill_manual(values=c("#ffc000","#002060"))+
  coord_flip()+
  custom_theme()+
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    legend.title = element_blank(),
    legend.position = "bottom"
  )
```

### **Q4 - Which part of Nairobi has the highest concentration of stores?**
```{r map of store location, echo=TRUE}
#making a leaflet map to understand geographical distribution of stores

leaflet(width = "100%") %>%
   # add a default basemap
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>%
  addMarkers(
    data = sales_revenue_data,
    lng = ~lon,
    lat = ~lat,
    # create custom labels
    label = paste(
      "Shop Name:",
      sales_revenue_data$shop_name, "<br>",
      "Shop Location:",
     sales_revenue_data$shop_location, "<br>",
      "Total revenue:",
      sales_revenue_data$total_revenue, "<br>",
      "Total sales quantity: ",
      sales_revenue_data$total_sales
    )
    %>%
      lapply(htmltools::HTML)
  )

```

Calculating the total store count per neighborhood to determine the areas with highest concentration
```{r store count per neighborhood, echo=TRUE}

#calculating the store count
sales_store_count <- sales_revenue_data %>%
  group_by(shop_location) %>%
  count(shop_name) %>%
  summarise(
    shop_count = sum(n)
  ) 
#sorting the data
top_salesstore_count <- sales_store_count %>% 
  arrange(desc(shop_count)) %>% top_n(5)

#inspecting data 
head(top_salesstore_count)

```

Plotting areas with the highest concentration of stores
```{r}
top_salesstore_count %>%  ggplot(aes(x=reorder(shop_location,shop_count),y=shop_count))+
  geom_col(fill="#ffc107")+
  labs(title ="Parts of Nairobi with the highest concentration of stores",y="",x="Neighborhood")+
  geom_text(aes(label=shop_count),color="black",
            hjust=0.6,vjust=0.5,family=f1,size=2.8)+
  coord_flip()+
  custom_theme()+
  theme(
  axis.text.x = element_blank()
  )
```

